---
title: "Exercise 8"
output: html_document
date: "2025-03-05"
---
For this exercise, we will practices testing for significant differences between two categorical variables. The procedure involves first generating a two-way frequency table, which we worked on in exercise 4. Then we will conduct a chi-square test on that frequency table to see if there are significant differences across the levels of that frequency table.

1. If you need to download the NHANES data base and code the nominal and ordinal variables as factors, please run this code chunk below:
```{r}
#Use this code chunk to download the NHANES 2017-2018 database and to label the nominal and ordinal variables as factor variables (if necessary):
NHANESdata <- read.csv("https://raw.githubusercontent.com/malenamalka/ANTH241-SP25/refs/heads/main/data/NHANES2017-28")
factor(c(NHANESdata$Gender, NHANESdata$Race, NHANESdata$HealthIns))
factor(c(NHANESdata$Education,NHANESdata$HHEducation,NHANESdata$HHIncome,NHANESdata$FeelBad,NHANESdata$HHFoodSec),order=TRUE)
factor(NHANESdata$BMIcat, order =TRUE, levels = c("Underweight","Healthy","Overweight","Obese"))
head(NHANESdata)
```

2. Load the following packages that you will use in this exercise.
```{r}
library(dplyr)
library(tidyr)
```

3. Let's examine the relationship between BMI categories and household food security:
```{r}
#a. Generate a two-way frequency table of BMIcat and HHFoodSec:
#table_name <- xtabs(~ CatVar1 + CatVar2, data=)
#print the table
#show the table in percentages: prop.table(table_name)*100
BMIcat_FS_table <- xtabs(~ BMIcat + HHFoodSec, data=NHANESdata)
BMIcat_FS_table
prop.table(BMIcat_FS_table)*100
#b. run the chi-square test:
#chisq.test(table_name)
chisq.test(BMIcat_FS_table)
```

4. Compare Gender and Race:
```{r}
#a. generate a frequency table by raw counts and percentages
Gender_race_table <- xtabs(~ Gender + Race, data=NHANESdata)
Gender_race_table
prop.table(Gender_race_table)*100
#b. run the Chi-square test on that table.
chisq.test(Gender_race_table)
```

Let's compare BMIcat by Education level
```{r}
BMIcat_Education_table <- xtabs(~ BMIcat + Education, data=NHANESdata)
BMIcat_Education_table
prop.table(BMIcat_Education_table)*100
chisq.test(BMIcat_Education_table)
```

