---
title: "Exercise 9"
output: html_document
date: "2025-03-19"
---
For this exercise, you will practice creating correlations between two continuous variables and illustrating these relationships as scatter plots.

1. You should download the NHANES data and label the nominal and ordinal variables as factors if you need to.
```{r}
#Use this code chunk to download the NHANES 2017-2018 database and to label the nominal and ordinal variables as factor variables (if necessary):
NHANESdata <- read.csv("https://raw.githubusercontent.com/malenamalka/ANTH241-SP25/refs/heads/main/data/NHANES2017-28")
factor(c(NHANESdata$Gender, NHANESdata$Race, NHANESdata$HealthIns))
factor(c(NHANESdata$Education,NHANESdata$HHEducation,NHANESdata$HHIncome,NHANESdata$FeelBad,NHANESdata$HHFoodSec),order=TRUE)
factor(NHANESdata$BMIcat, order =TRUE, levels = c("Underweight","Healthy","Overweight","Obese"))
head(NHANESdata)
```

2. The following code opens the packages that you will need for this exercise.
```{r}
library(dplyr)
library(ggplot2)
```

3. Some of the continuous variables contain cells that code for "don't know" or "refused" (e.g, 7777s and 9999s). The code chunk below removes these cells.
```{r}
NHANESdata_reduced <- NHANESdata%>%
  filter(SMSpend<4286, Vigorous<481, Moderate<541, Sedentary<1321)
head(NHANESdata_reduced)
```

4. Create a scatterplot of BMI and Waist Circumference and generate a Pearson correlation between these two variables:
a. What is the null hypothesis?
b. What is the alternative hypothesis?
```{r}
#scatter plot code: #ggplot(aes(x=DepVar, y=IndVar), data=) + geom_point(color="") + geom_smooth(method=lm, color="") + labs(x="", y="") + theme_bw()

#correlation test code: cor.test(data$DepVar, data$IndVar, method="pearson")

```
Questions: What is the Pearson correlation coefficient of BMI and Waist Circumference? Is this correlation coefficient significant? Finally, what do these statistics indicate about the null/alternative hypothesis?

5. Generate a scatter plot and Pearson correlation test of the relationship between BMI and daily sugar intake (DR1TSUGR)
a. What is the null hypothesis?
b. What is the alternative hypothesis?
```{r}
#scatter plot code: #ggplot(aes(x=DepVar, y=IndVar), data=) + geom_point(color="") + geom_smooth(method=lm, color="") + labs(x="", y="") + theme_bw()

#correlation test code: cor.test(data$DepVar, data$IndVar, method="pearson")

```
Question: What is the Pearson correlation coefficient of BMI and daily sugar intake? Is this correlation coefficient significant? Finally, what do these statistics indicate about the null/alternative hypothesis? 

6. Generate a scatter plot and correlation test of BMI and daily fat intake (DR1TTFAT)?
a. What is the null hypothesis?
b. What is the alternative hypothesis?
```{r}
#scatter plot code: #ggplot(aes(x=DepVar, y=IndVar), data=) + geom_point(color="") + geom_smooth(method=lm, color="") + labs(x="", y="") + theme_bw()

#correlation test code: cor.test(data$DepVar, data$IndVar, method="pearson")

```
Question: What is the Pearson correlation coefficient of BMI and daily fat intake? Is this correlation coefficient significant? Finally, what do these statistics indicate about the null/alternative hypothesis?

7. Generate a scatter plot and correlation test of the relationship between BMI and moderate daily exercise:
a. What is the null hypothesis?
b. What is the alternative hypothesis?
```{r}
#scatter plot code: #ggplot(aes(x=DepVar, y=IndVar), data=) + geom_point(color="") + geom_smooth(method=lm, color="") + labs(x="", y="") + theme_bw()

#correlation test code: cor.test(data$DepVar, data$IndVar, method="pearson")

```
Question: What is the Pearson correlation coefficient of BMI and moderate daily exercise? Is this correlation coefficient significant? Finally, what do these statistics indicate about the null/alternative hypothesis?

7. Generate a scatter plot and Pearson correlation test for the relationship between BMI and daily sedentary minutes.
a. What is the null hypothesis?
b. What is the alternative hypothesis?
```{r}
#scatter plot code: #ggplot(aes(x=DepVar, y=IndVar), data=) + geom_point(color="") + geom_smooth(method=lm, color="") + labs(x="", y="") + theme_bw()

#correlation test code: cor.test(data$DepVar, data$IndVar, method="pearson")

```
Question: What is the Pearson correlation coefficient of BMI and daily sedentary minutes? Is this correlation coefficient significant? Finally, what do these statistics indicate about the null/alternative hypothesis?

8. Generate a scatter plot and Pearson correlation test of the relationship between BMI and money spent at the grocery store (SMSpend).
a. What is the null hypothesis?
b. What is the alternative hypothesis?
```{r}
#scatter plot code: #ggplot(aes(x=DepVar, y=IndVar), data=) + geom_point(color="") + geom_smooth(method=lm, color="") + labs(x="", y="") + theme_bw()

#correlation test code: cor.test(data$DepVar, data$IndVar, method="pearson")

```
Question: What is the Pearson correlation coefficient of BMI and money spent at the grocery store? Is this correlation coefficient significant? Finally, what do these statistics indicate about the null/alternative hypothesis?

