---
title: "Exercise 10"
output: html_document
date: "2025-03-23"
---
For exercise 10, you will practice generating OLS regression models that predict the outcome of a continuous dependent variable from a continuous independent variable.

1. If necessary, download the NHANES data set and label the ordinal and nominal variables as factor variables.

```{r}
#Use this code chunk to download the NHANES 2017-2018 database and to label the nominal and ordinal variables as factor variables (if necessary):
NHANESdata <- read.csv("https://raw.githubusercontent.com/malenamalka/ANTH241-SP25/refs/heads/main/data/NHANES2017-28")
factor(c(NHANESdata$Gender, NHANESdata$Race, NHANESdata$HealthIns))
factor(c(NHANESdata$Education,NHANESdata$HHEducation,NHANESdata$HHIncome,NHANESdata$FeelBad,NHANESdata$HHFoodSec),order=TRUE)
factor(NHANESdata$BMIcat, order =TRUE, levels = c("Underweight","Healthy","Overweight","Obese"))
head(NHANESdata)
```

2. If necessary, use the code chunk below to remove the cells with 7777s and 9999s, etc. from some of the continuous variables.
```{r}
NHANESdata_reduced <- NHANESdata%>%
  filter(SMSpend<4286, Vigorous<481, Moderate<541, Sedentary<1321)
head(NHANESdata_reduced)
```

3. Generate a linear regression model that predicts BMI (depedent variable) from waist circumference (independent variable).
```{r}
#model_name <- lm(DepVar~IndVar, data=)
#summary(model_name)
```
What is the regression equation for this model?

What is the multiple R-squared value for this equation?

What is the p-value?

What do these results indicate about the relationship between BMI and waist circumference?

4. Generate a linear regression model that predicts BMI from RFITP. Insert a code chunk to create this regression model.


Questions: What is the linear regression equation of this model? 
What is the multiple R-squared value? 
What is the p-value of the model? 
What do these results indicate about the relationship between BMI and RFITP?


5. Generate a linear regression model that predicts BMI from daily caloric intake (DR1TKCAL). After you create and run the model in a new code chunk below, answer the same questions as above. 


6. Do the same steps as above for BMI predicted from daily carbohydrate intake.


7. Do the same steps as above for BMI predicted from daily sugar intake.


8. Do the same steps as above for BMI predicted from daily fat intake.


9. Do the same steps as above for BMI predicted from vigorous daily exercise.


10. Do the same steps as above for BMI predicted from sedentary minutes per day.


After generating these 10 regression models, what can you conclude about the relationship between being overweight or obese with dietary intake and physical activity?


