---
title: "Exercise 3_blank"
output: html_document
date: "2025-02-04"
---

For this exercise, we are following along with using R and RStudio in chapter 6 of S&B and chapter K, chapters 1 and 2. The goal of this exercise is to get you using R and R Studio by following the step-by-step codes in the examples from these two chapters.

The first step in using R and R Studio is to download the software to your computer:

Once you have installed the software to your computer, you can now follow along completing the exercises 6.1.3, 6.1.4, and 6.2 in the S&B textbook

Exercise 6.13
You will need to install and open several packages that allow us to complete specific tasks. Create a chunk if code via the Code tab, then "insert chunk". In the code chunk, add the prompts for install.packages and librarys as written on pp. 55-56.

```{r}
install.packages("tidyverse")
install.packages("dplyr")
install.packages("modelsummary")
install.packages("foreign")
library(tidyverse)
library(dplyr)
library(modelsummary)
library(foreign)
```

Exercise 6.1.4
```{r}
#a. Creating a vector "x" with the values, 1,2, and 3

#b. Creating a vector "y" with the values "yes" and "no"

#Create a data frame with the x and y vectors created above:

#Write "dat" below will display the data frame

#We can also single out the second column (vector 7) using the $ operator


```

To see what happens when you make an error, let's attempt to assign a value of variable W to variable Z
```{r}
z <- w
```
The reason for the error is that we assigned a value to a variable that does not exist.

Exercise 6.2: Importing data into R
a. Setting a working directory. I recommend setting your working directory to folders on your One Drive.
```{r}
#setwd("your file path")
getwd()
```

Now, we will import a dataset into R that the authors, S&B, use as examples. The data is called "PartyData.csv"
```{r}
#Running this code will download the data, "PartyData.csv" from a respository on the web.
party_data <- read.csv("https://raw.githubusercontent.com/nickbordeleau/QMSS/main/PartyData.csv")
View(party_data)
head(party_data)
#View(party-data) shows the data file in a new tab above.
#head(party_data) shows the first six rows of the data file in this RMD document.
```

Now we will complete some exercises in the Kabacoff book that introduce using R and R studio. Follow along in chapters 1 and 2 in that book. The exercises below are taken directly from those two chapters.
```{r}
#This is listing 1.1 from page 9 in Kabacoff, in which we create a database with two variables (age and weight)

```

Kabacoff, listing 2.1, page 24 on creating matrices:
```{r}
#1. Creating a 5x4 matrix

```

```{r}
#2: create a 2x2 matrix filled by rows

```

```{r}
#3. Create a 2x2 matrix filled by columns

```

Kabacoff, Listing 2.2 p. 24:
```{r}
#Using matrix subscripts
#First, create a 2x5 matrix containing numbers 1-10

#second, the elements in the second row are selected

#third, the elements in the second column are selected

#Fourth, the elements in the first and fourth column are selected

#finally, the elements in the first row and fourth and fifth columns are selected

```

Kabacoff, listing 2.3, p. 25
```{r}
#Creating a three-dimensional (2x3x4) array of numbers

```

Kabacoff, listing 2.4, p. 26
```{r}
#Creating a data frame

```

Kabacoff, listing 2.5, p. 26-27
```{r}
#Specifying elements of a data frame
#1. To specify the first two columns of the data frame, patientID and age

#2. To specify the columns, "diabetes" and "status":

#3, To show the column, "age"

```

Kabacoff, listing 2.6, pp. 29-30
```{r}
#Using factors:
#1. Enter data as vectors:

#the statement below indicates that diabetes is a nominal variable (factor variable)
diabetes <- factor(diabetes)

#the statement below indicates that status is an ordinal variable that R treats as a factor in an ordered sequence.
status <- factor(status, order=TRUE)

#2. Display the object structure to show that diabetes is a factor and status is an ordered factor

#3. Display the object summary to show minimum, maximum, mean, and quartiles for continuous variable age and frequency counts for categorical variables, diabetes and status

```

Kabacoff, Listing 2.7, p. 30
```{r}
#Creating a list - follow along the steps in this listing to create "My First List"
```
