---
title: "Exercise 12"
output: html_document
date: "2025-04-02"
---
In this exercise, we will learn odds ratios for comparing the odds of developing a particular condition based on different levels of a categorical (nominal or ordinal) variable. Odds ratios are commonly used in public health to understand the likelihood, or odds, of developing some kind of health condition. 

In odds ratios, the dependent variable is a dichotomous nominal variable. For example, whether you have health insurance (yes or no) is a dichotomous variable. Whether you have type 2 diabetes (yes or no) is another example of a dichotomous variable. If your data is not in a dichotomous nominal form, you can easily convert that variable with some simple code.

There are many ways to compute odds ratios in R Studio, many of which involve sophisticated applications of logistic regression models. Here, we will simplify by using an easily accessible function in the "epitools" package.

1. Download the NHANES 2017-2018 data set if necessary.
```{r}
#Use this code chunk to download the NHANES 2017-2018 database and to label the nominal and ordinal variables as factor variables:
NHANESdata <- read.csv("https://raw.githubusercontent.com/malenamalka/ANTH241-SP25/refs/heads/main/data/NHANES2017-28")
factor(c(NHANESdata$Gender, NHANESdata$Race, NHANESdata$HealthIns))
factor(c(NHANESdata$Education,NHANESdata$HHEducation,NHANESdata$HHIncome,NHANESdata$FeelBad,NHANESdata$HHFoodSec),order=TRUE)
factor(NHANESdata$BMIcat, order =TRUE, levels = c("Underweight","Healthy","Overweight","Obese"))
```

2. Load the packages that we will use for this exercise:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
install.packages("epitools")
library(epitools)
install.packages("ggmosaic")
library(ggmosaic)
```

3. Use the code below to create a data set without the "777s" etc. from the continuous variables. I include "HealthIns" in this code chunk too.
```{r}
NHANESdata_reduced <- NHANESdata%>%
  filter(SMSpend<4286, Vigorous<481, Moderate<541, Sedentary<1321, HealthIns<3)
```

4. The health insurance variable is currently coded as 1=yes (have health insurance) and 2=no (no health insurance). I want to change these responses to "Yes" and "No" , and I want "No" to be sorted first before "Yes". To accomplish this, run the code below:
```{r}
NHANESdata_reduced$HealthInsYN[NHANESdata_reduced$HealthIns==1] <- "Yes"
NHANESdata_reduced$HealthInsYN[NHANESdata_reduced$HealthIns==2] <- "No"
factor(c(NHANESdata_reduced$HealthInsYN, levels=c("No", "Yes")))
head(NHANESdata_reduced)
```

5. Calculate the odds ratio of having health insurance (dependent variable) by race (independent variable)
```{r}
#1. Create the odds table:
#odds_table_name <- table(data$IndVar, data$DepVar)

#2. Run the oddsratio on the table you created:
#oddsratio(odds_table_name)

#3. Graph the results in a mosaic table:
#ggplot(data=) + geom_mosaic(aes(x=product(IndVar), fill=DepVar), na.rm=TRUE, show.legend=FALSE) + labs(title="", x="", y="")

```
Questions: Compared to Mexican Americans, what are the odds ratios having health insurance for each racial category?

Are these odds ratios significant?

6. Calculate the odds ratio of having health insurance by BMI category:
```{r}

```
Questions: 
Compared to the healthy level, what are the odds ratio of having health insurance of the other three BMI categories?

Are any of these odds ratios significant?


7a. Calculate odds ratio of moderate exercise minutes by BMI category. Odds ratios do not work on continuous variables. We have to convert moderate exercise into a dichotomous nominal variable. Here, we will create a new variable, "mod_exerc", in which "Little" = 30 minutes or less of moderate exercise, and "More" = more than 30 minutes of exercise.
```{r}

```

7b. Now, calculate the odds ratio of mod_exerc by BMIcat
```{r}

```
Questions: 
Compared to the healthy level, what are the odds ratios of "More" exercise minutes for the other three BMI categories?

Are these odds ratios statistically significant?

8. Let's calculate a new odds ratio.
Independent variable:
Dependent variable: