---
title: "Exercise 6"
output: html_document
date: "2025-02-24"
---

For this exercise, you will generate graphs that show univariate, descriptive statistics for nominal, ordinal, and continuous variables.
```{r}
#Use this code chunk to download the NHANES 2017-2018 database and to label the nominal and ordinal variables as factor variables (if necessary):
NHANESdata <- read.csv("https://raw.githubusercontent.com/malenamalka/ANTH241-SP25/refs/heads/main/data/NHANES2017-28")
factor(c(NHANESdata$Gender, NHANESdata$Race, NHANESdata$HealthIns))
factor(c(NHANESdata$Education,NHANESdata$HHEducation,NHANESdata$HHIncome,NHANESdata$FeelBad,NHANESdata$HHFoodSec),order=TRUE)
factor(NHANESdata$BMIcat, order =TRUE, levels = c("Underweight","Healthy","Overweight","Obese"))
head(NHANESdata)
```

1. Load the following packages to complete this exercise: ggplot2, tidyr, and dplyr.
```{r}
install.packages("ggplot2")
library(ggplot2)
library(tidyr)
library(dplyr)
```

1. Bar charts illustrate variation for nominal and ordinal variables.
```{r}
#a. Bar chart for a nominal variable showing raw counts: Create the bar chart for BMI categories (BMIcat)
#ggplot (data, aes(x=Catvar)) + geom_bar() + labs(title="", x="", y="")
ggplot(NHANESdata,aes(BMIcat)) +geom_bar(fill="red") + labs(title = "Frequency of BMI cat", x = "BMIcat", y="count")
```

```{r}
#b. Bar chart showing BMI categories by percentages:
#new_table_name <- with(NHANESdata, table(BMIcat))
BMIcat_table <- with(NHANESdata, table(BMIcat))
#turn the above table into proportions: prop.table(new_table_name)
prop.table(BMIcat_table)
#turn the table into percentages: prop.table(new_table_name)*100
#new_data_frame <- as.data.frame(prop.table(new_table_name)*100
BMIcat_barchart <- as.data.frame(prop.table(BMIcat_table)*100)
BMIcat_barchart
#ggplot(new_data_frame, aes(BMIcat, Freq)) + geom_col()
ggplot(BMIcat_barchart, aes(BMIcat,Freq)) +geom_col(fill="turquoise")
```

c. Construct bar charts for one ordinal variable in the code chunk below:
```{r}

```

d. Constructing bar charts for a nominal/ordinal variable grouped by another nominal/ordinal variable:
```{r}
#BMI_cats by gender: stacked, grouped and filled bar charts:
#table(data$Independent variable, data$Dependent variable)
table(NHANESdata$Gender, NHANESdata$BMIcat)
#Stacked bar chart:
#ggplot(data, aes(x=Dependent variable, fill=Independent variable)) + geom_bar(position = "stack") + labs(title="", x="", y="")
ggplot(NHANESdata,aes(x=BMIcat, fill=Gender)) +geom_bar(position = "stack") + labs(title="BMI categories by gender", x="Gender", y="Frequency")
#grouped bar chart:
#ggplot(data, aes(x=Dependent variable, fill=Independent variable)) + geom_bar(position = "dodge") + labs(title="", x="Gender", y="Frequency")
ggplot(NHANESdata,aes(x=BMIcat, fill=Gender)) +geom_bar(position = "dodge") + labs(title="BMI categories by gender", x="Gender", y="Frequency")
```
```{r}
ggplot(NHANESdata,aes(x=BMIcat, fill=Gender)) +geom_bar(position = "stack") + facet_wrap(~Race) + labs(title="BMI categories by gender", x="Gender", y="Frequency")
ggplot(NHANESdata,aes(x=BMIcat, fill=Gender)) +geom_bar(position = "stack") + facet_grid(~Race) + labs(title="BMI categories by gender", x="Gender", y="Frequency")
```


e. Constructing bar charts for mean +/- standard error of continuous variables
```{r}
#1 generate means and standard errors of BMI(dependent var.) by gender (independent var)
#new_data_frame <- data%>%
#drop_na(Dependent variable)%>%
#group_by(Independent variable)
#summarize(n=n(), mean=mean(dependent variable), sd=sd(dependent variable))
#to print new data frame: new_data_frame

#2 Plot means in a sorted bar chart and adds error bars that represent the standard deviation
#ggplot(new_data_frame, aes(x=reorder(Independent variable, mean), y=mean)) + geom_bar(stat="identity", fill="color") + geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2)
#labs(x="", y="", title="", subtitle="")

BMI_gender <- NHANESdata %>% 
  drop_na(BMI) %>%
  group_by(Gender) %>% 
  summarize(count=n(), mean = mean(BMI),sd = sd(BMI))
BMI_gender

ggplot(BMI_gender, aes(x=reorder(Gender, mean), y=mean)) + 
  geom_bar(stat="identity", fill = "purple") +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2) +
  labs(y = "", 
       x="", 
       main = ""
  )
#labs(x="", y="", title="", subtitle="")

```

2. Histograms illustrate variation in continuous variables:
```{r}
#a. simple histogram: ggplot(data, aes(x=contvar)) + geom_histogram() + labs(title="", x="", y="")
#construct a histogram for BMI:
ggplot(NHANESdata, aes(x=BMI)) + geom_histogram(fill = "navy") + labs(title = "Colored histograms with 20 bins", x="BMI", y="count")
```

```{r}
#b. colored histogram with 20 bins: ggplot(data, aes(x=contvar)) + geom_histogram(bins=20, color="white", fill="steelblue") + labs(title="Colored histogram with 20 bins", x="", y="")
#do the above code for BMI:
ggplot(NHANESdata, aes(x=BMI)) + geom_histogram(bins = 30, color = "white", fill = "steelblue") + labs(title = "Colored histograms with 30 bins", x="BMI", y="count")
```

```{r}
#c. Histogram with percentages:
#ggplot(data, aes(x=contvar, y=..density..)) + geom_histogram(bins=20, color="white", fill="steelblue") + scale_y_continuous(labels=scales::percent) + labs(title="", y="percent", x="")
ggplot(NHANESdata,aes(x=BMI,y=..density..)) +geom_histogram()
```

d: Create a histogram of a different continuous variable:

3. Box plots: a box and whisker plot describes the distribution of a continuous variable by plotting its five-number summary: minimum, lower quartile (25th percentile), median (50th percentile), upper quartile (75th percentile), and maximum, and the outliers (values outside of the +/- 1.5xQR)
```{r}
#a.ggplot(data, aes(x="", y=Contvar)) + geom_boxplot() + labs(y="", x="", title="")
#Create a box plot of the amount of protein consumed in a day (DR1TPROT)
ggplot(NHANESdata,aes(y=DR1TPROT)) +geom_boxplot()
```


```{r}
#b. Create a box plot that comnpares a continuous variable across different levels of a nominal or ordinal variable: DR1TPROT by race:
#ggplot(data, aes(x=Independent variable, y=Contvar)) + geom_boxplot(fill="color") + labs(x="Independent variable", y="Contvar", title="")

```

