---
title: "Exercise 3_blank"
output: html_document
date: "2025-02-04"
---

For this exercise, we are following along with using R and RStudio in chapter 6 of S&B and chapter K, chapters 1 and 2. The goal of this exercise is to get you using R and R Studio by following the step-by-step codes in the examples from these two chapters.

The first step in using R and R Studio is to download the software to your computer:

Once you have installed the software to your computer, you can now follow along completing the exercises 6.1.3, 6.1.4, and 6.2 in the S&B textbook

Exercise 6.13
You will need to install and open several packages that allow us to complete specific tasks. Create a chunk if code via the Code tab, then "insert chunk". In the code chunk, add the prompts for install.packages and librarys as written on pp. 55-56.

```{r}
install.packages("tidyverse")
install.packages("dplyr")
install.packages("modelsummary")
install.packages("foreign")
library(tidyverse)
library(dplyr)
library(modelsummary)
library(foreign)
```

Exercise 6.1.4
```{r}
#a. Creating a vector "x" with the values, 1,2, and 3
x <- c(1,2,3)
#b. Creating a vector "y" with the values "yes", "no" and "maybe"
y <- c("yes", "no", "maybe")
#Create a data frame with the x and y vectors created above:
dat <- data.frame(x,y)
#Write "dat" below will display the data frame
dat
#We can also single out the second column (vector 7) using the $ operator
dat$y
```

To see what happens when you make an error, let's attempt to assign a value of variable W to variable Z
```{r}
z <- w
```
The reason for the error is that we assigned a value to a variable that does not exist.

Exercise 6.2: Importing data into R
a. Setting a working directory. I recommend setting your working directory to folders on your One Drive.
```{r}
#setwd("your file path")
getwd()
```

Now, we will import a dataset into R that the authors, S&B, use as examples. The data is called "PartyData.csv"
```{r}
#Running this code will download the data, "PartyData.csv" from a respository on the web.
party_data <- read.csv("https://raw.githubusercontent.com/nickbordeleau/QMSS/main/PartyData.csv")
View(party_data)
head(party_data)
#View(party-data) shows the data file in a new tab above.
#head(party_data) shows the first six rows of the data file in this RMD document.
```

Now we will complete some exercises in the Kabacoff book that introduce using R and R studio. Follow along in chapters 1 and 2 in that book. The exercises below are taken directly from those two chapters.
```{r}
#This is listing 1.1 from page 9 in Kabacoff, in which we create a database with two variables (age and weight)
age <- c(1,3,5,2,11,9,3,9,12,3)
weight <- c(4.4, 5.3, 7.2, 5.2, 8.5, 7.3, 6.0, 10.4, 10.2, 6.1)
mean(weight)
sd(weight)
cor(age,weight)
plot(age,weight)
```

Kabacoff, listing 2.1, page 24 on creating matrices:
```{r}
#1. Creating a 5x4 matrix
y <- matrix(1:20, nrow=5, ncol=4)
y
cells <- c(1,26,24,68)
rnames <- c("R1", "R2")
cnames <- c("C1", "C2")

```

```{r}
#2: create a 2x2 matrix filled by rows
mymatrix <- matrix(cells, nrow=2, ncol=2, byrow=TRUE,
                   dimnames=list(rnames, cnames))
mymatrix
```

```{r}
#Create a 2x2 matrix filled by columns
mymatrix <- matrix(cells, nrow=2, ncol=2, byrow=FALSE,
                   dimnames=list(rnames, cnames))
mymatrix
```

Kabacoff, Listing 2.2 p. 24:
```{r}
#Using matrix subscripts
#First, create a 2x5 matrix containing numbers 1-10
x <- matrix(1:10, nrow=2)
x
#second, the elements in the second row are selected
x[2,]
#third, the elements in the second column are selected
x[,2]
#Fourth, the elements in the first and fourth column are selected
x[1,4]
#finally, the elements in the first row and fourth and fifth columns are selected
x[1, c(4,5)]
```

Kabacoff, listing 2.3, p. 25
```{r}
#Creating a three-dimensional (2x3x4) array of numbers
dim1 <- c("A1", "A2")
dim2 <- c("B1", "B2", "B3")
dim3 <- c("C1", "C2", "C3", "C4")
z <- array(1:24, c(2,3,4), dimnames=list(dim1, dim2, dim3))
z
```

Kabacoff, listing 2.4, p. 26
```{r}
#Creating a data frame
patientID <- c(1, 2, 3, 4)
age <- c(25,34,28,52)
diabetes <- c("Type1", "Type2", "Type1", "Type1")
status <- c("Poor", "Improved", "Excellent", "Poor")
patientdata <- data.frame(patientID, age, diabetes, status)
patientdata
```

Kabacoff, listing 2.5, p. 26-27
```{r}
#Specifying elements of a data frame
#1. To specify the first two columns of the data frame, patientID and age
patientdata[1:2]
#2. To specify the columns, diabetes and status:
patientdata[c("diabetes", "status")]
#3, To show the column, age
patientdata$age
```

Kabacoff, listing 2.6, pp. 29-30
```{r}
#Using factors:
#1. Enter data as vectors:
patientID <- c(1, 2, 3, 4)
age <- c(25, 34, 28, 52)
diabetes <- c("Type1", "Type2", "Type1", "Type1")
status <- c("Poor", "Improved", "Excellent", "Poor")
diabetes <- factor(diabetes)
#the above statement indicates that diabetes is a nominal variable (factor variable)
status <- factor(status, order=TRUE)
#the above statement indicates that status is an ordinal variable that R treats as a factor in an ordered sequence.
patientdata <- data.frame(patientID, age, diabetes, status)
#2. Displays the object structure to show that diabetes is a factor and status is an ordered factor
str(patientdata)
#3. Displays the object summary to show minimum, maximum, mean, and quartiles for continuous variable age, and it also shows frequency counts for categorical variables, diabetes and status
summary(patientdata)
```

Kabacoff, Listing 2.7, p. 30
```{r}
#Creating a list - follow along the steps in this listing to create "My First List"
```
