---
title: "Feb 11 class exercise"
output: html_document
date: "2025-02-10"
---
For this exercise, we will:
1. practice downloading the RMD file from GitHub
2. download the NHANES 2017-2018 database
3. convert the nominal and ordinal variables in this database to be read accordingly in R Studo
4. upload our completed RMD file to GitHub

2. Once you have downloaded the RMD file from GitHub, we will download the raw data for the NHANES 2017-2018 dataset that I created to work in R Studio.

```{r}
#create a name for your data set. Below, I call mine "NHANESdata": Go to the class github site, click on "data", then "2017-28 data", then click on "raw", then copy the URL address into the parentheses below with "" marks
NHANESdata <- read.csv()
#to show the first six rows of the data set:
head(NHANESdata)
```

3. What are the nominal/categorical variables in this data set?

Let's tell R Studio to treat these three variables as categorical ones:
```{r}
factor()
factor()
factor()
```

What are the ordinal variables in this data set?
Let's tell R Studio to treat these six variables as ordinal ones:
```{r}
factor()
```

4. Let's upload this RMD file to github
